Poincaré and SimBio: a versatile and extensible Python ecosystem for modeling systems.

### 2024-04-04 article.tex
Add unsupported SBML features
Mention independent variable

Poincare is a package to define and simulate dynamical systems.
Based on Python inmutable dataclasses,
it provides a \texttt{System} class,
where one can define
\texttt{Constant}s; which can be numbers or refer to other constants;
\texttt{Parameter}s, which can be numbers or time-dependent expressions;
\texttt{Variable}s, which represent the state of the system and must be provided with an initial condition;
and create equations linking a variable's derivative with an expression (Figure~\ref{fig-first-order}).
It also allows to define higher-order systems
by assigning an initial condition to a \texttt{Derivative} (Figure~\ref{fig-second-order}).

@@

Several commonly used reactions are predefined as \texttt{MassAction} subclasses,
such as \texttt{MichaelisMenten} (\(S + E \leftrightarrow ES \rightarrow P + E\))
and its approximate form without the intermediate species \(ES\),
and it is also simple to implement used-defined ones
as subclasses of \texttt{RateLaw} or \texttt{MassAction}.
Additionally, SimBio supports importing models from \ac{SBML},
and downloading them directly from BioModels \cite{malik-sheriffBioModels15Years2020} (Figure~\ref{fig-simbio-io}).

>>>>

Poincare is a package to define and simulate dynamical systems.
Based on Python inmutable dataclasses,
it provides a \texttt{System} class,
where one can define
\texttt{Constant}s; which can be numbers or refer to other constants;
\texttt{Parameter}s, which can be numbers or time-dependent expressions;
\texttt{Variable}s, which represent the state of the system and must be provided with an initial condition;
and create equations linking a variable's derivative with an expression (Figure~\ref{fig-first-order}).
It also allows to
create an \texttt{Independent} variable to define non-autonomous systems,
and define higher-order systems by assigning an initial condition to a \texttt{Derivative} (Figure~\ref{fig-second-order}).

@@

Several commonly used reactions are predefined as \texttt{MassAction} subclasses,
such as \texttt{MichaelisMenten} (\(S + E \leftrightarrow ES \rightarrow P + E\))
and its approximate form without the intermediate species \(ES\),
and it is also simple to implement used-defined ones
as subclasses of \texttt{RateLaw} or \texttt{MassAction}.
Additionally, SimBio supports importing models from \ac{SBML},
and downloading them directly from BioModels \cite{malik-sheriffBioModels15Years2020} (Figure~\ref{fig-simbio-io}).
Work is in progress to support exporting to \ac{SBML} and add more \ac{SBML} features.
Currently \ac{SBML} unsupported features include
algebraic rules, constraints, events, reactions with the fast attribute or with math stoichiometry,
units and compartments with size different from 1.
